\outline{1} {Appendix A Henkin's Proof Outline}
\centerline{\xmplbxi APPENDIX A}
\medskip
\centerline{\xmplbx HENKINS'S PROOF OUTLINE}
\bigskip

\item{1.} Reduce the characterization of $T[\Gamma]$ to that of $T$ (\jumplink{4.1.reduction}{Reduction Theorem}).
\item{2.} Reduce the validity form of the \jumplink{4.2.completeness_form1}{Completeness Theorem} to the consistency form.
\item{3.} Show that the restriction of a model $\SA' | L(T)$ is a model of $T$; so if we
obtain a model for an extension, we also obtain a model for the base theory.
\item{4.} Given a consistent theory $T$ we need to find a model via syntactical means.
\item{5.} Let the theorems of $T$ say what is truth of the variable-free terms, which
denote individuals.
\item{6.} Introduce the canonical structure $\SA$ (also called the {\it term structure})
using the equivalence classes of variable-free terms as individuals.
\item{7.} Show that the atomic closed formulas are valid in the canonical structure
iff they are theorems.
\item{8.} The two obstacles for showing that the same applies to general closed formulas
are:
\itemitem{(i)} There might be not enough variable-free terms.
\itemitem{(ii)} There might be formulas which are undecidable.
\item{9.} To handle (i), Henkin theories are introduced which guarantee that a constant
exists for every closed instantiation (also called the {\it witness property}).
\item{10.} Show that if $T$ is a complete Henkin theory, then the canonical structure
for $T$, is a model of $T$.
\item{11.} Show that we can construct a Henkin theory $T_c$ from a consistent theory $T$
using the special constants and axioms, and that $T_c$ is a conservative extension of $T$.
\item{12.} Show that every consistent theory has a simple complete extension (\jumplink{4.2.lindenbaum}{Lindenbaum's Theorem}).

\vfill 

\bigskip
\hrule

\centerline{
\xymatrix@C=10pc@R=4pc{
\hbox{\bf Language} & \hbox{\bf Theory} & \hbox{\bf Structure} \\
L\ar[d]^{\hbox{extension}}\ar[r] & {T\hbox{(consistent)}}\ar[d]^{\hbox{\jumplink{4.2.lemma3}{conservative extension}}} & {\SA | L}\ar[l]_{\hbox{\jumplink{4.2.lemma1}{model}}} \\
L_c\ar[r] & {T_c\hbox{(consistent, Henkin)}}\ar[d]^{\hbox{\jumplink{4.2.lindenbaum}{simple extension}}} \\
L_c\ar[r] & {U\hbox{(complete, Henkin)}} & {\SA \hbox{(canonical)}}\ar[l]_{\hbox{\jumplink{4.2.lemma2_cor}{model}}}\ar[uu]_{\hbox{restriction}} }
}

\vfill
\break
